Reconstructing Propositional Proofs in Type Theory

### 2017-10-27 paper/sections/type-theory.tex
[ paper ] addtions in type theory.

To reconstruct proofs generated by \Metis, we have formalized in type
theory the propositional logic (we briefly describe classical propositional logic in Appendix~\ref{app:CPL}).

To accomplish that formalization, the reader must take into account,
type theory and intuitionistic logic are based on the same basis of
the spirit of constructive mathematics. Therefore, proving propositions
demands a construction of the proof but in classical logic, they may
no exist and some proofs are proved by refutation. That is, using
a non-constructive method.

Therefore, to reconstruct \Metis derivations, we included as an axiom, the
principle of excluded third (\abbre{PEM}) since \Metis proves by refutation.
Assuming \abbre{PEM}, we can justify refutation proofs by deriving from it the
\emph{reductio ad absurdum} rule (\abbre{RAA}). The \abbrev{RAA} rule is the
formulation of the principle of proof by contradiction, that is, a derivation of
a contradiction, $⊥$, from the hypothesis $(¬ φ)$, is a derivation of $φ$.

\begin{notation}
We write theorems as the family of dependent types, $Γ ⊢ φ$ where
the turnstile symbol (⊢) represents the syntactical consequence relation
between a set of premises, $Γ : \List~\Prop$ and a conclusion $φ : \Prop$.
$\abbre{List}$ type is the usual inductive type for lists.
\end{notation}

We implemented in~\cite{AgdaProp} the syntactical consequence relation in a
similar way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken} and formation rules for
connectives as the constructors for new theorems in
Fig.~\ref{fig:CPL-inference-rules}.

>>>>

Type theory and intuitionistic logic are accompanying theories since they are
based on the same basis of constructive mathematics. Proving propositions in
these theories demands a witness construction for the proof. Nonetheless,
classical logic does not always have constructive proofs and some of its proofs
are proved by non-constructive methods like refutation.

To reconstruct proofs generated by \Metis, we have formalized in type theory the classical propositional logic.
A  briefly description of it is presented in Appendix~\ref{app:CPL}. In this
formalization, we assume the principle of excluded third (\abbre{PEM}) as the
only axiom since \Metis is a prover for classical logic. Assuming \abbre{PEM},
we can justify refutation proofs by deriving from it the \emph{reductio ad
absurdum} rule (\abbre{RAA}). The \abbrev{RAA} rule is the formulation of the
principle of proof by contradiction, that is, a derivation of a contradiction,
$⊥$, from the hypothesis $(¬ φ)$, is a derivation of $φ$.

\begin{notation}
We write theorems in \CPL as the family of dependent types, $Γ ⊢ φ$ where
the turnstile symbol (⊢) represents the syntactical consequence relation
between a set of premises, $Γ : \List~\Prop$ and a conclusion $φ : \Prop$.
$\abbre{List}$ type is the usual inductive type for lists.
\end{notation}

We implemented in~\cite{AgdaProp} the syntactical consequence relation in a
similar way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken}, formation and elimination
rules for connectives and the PEM axiom as the constructors of the theorem type
in Fig.~\ref{fig:CPL-inference-rules}.
